---
title: "Grocery Sales Forecast"
author:
  Christine Vu^[University of San Diego, cvu@sandiego.edu], Dave Friesen^[University of San Diego, dfriesen@sandiego.edu]
date: "12/12/2022"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output:
  html_document:
    css: "style.css"
  pdf_document: default
---

<style>
.main-container {
  max-width: 1024px;
}
</style>


### Objective and Hypothesis

#### [. . .]


```{r setup, echo = FALSE, message = FALSE}
# Load R libraries

# Expand output width and minimize exp notation
options(width = 150)
options(scipen = 100)
options(digits = 1)

# Set style defaults
knitr::opts_chunk$set(class.source = "source")
knitr::opts_chunk$set(class.output = "output")
knitr::opts_chunk$set(fig.width = 10, fig.height = (10 * .45), fig.align = "center")
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NA)
```

### Data Load and Validation

```{r data_load_validation}
# Load dataset(s); assumes folder structure with data parallel to src
train_df <- read.csv("../data/train.csv", header = TRUE)
test_df <- read.csv("../data/test.csv", header = TRUE)
stores_df <- read.csv("../data/stores.csv", header = TRUE)
oil_df <- read.csv("../data/oil.csv", header = TRUE)
events_df <- read.csv("../data/holidays_events.csv", header = TRUE)

# Data validation and understanding, including structure, content, and statistical characteristics covered below
```


###### Data Structure Review

```{r data_structure_review}
# Summarize base datasets
str(train_df)
str(test_df)
str(stores_df)
str(oil_df)
str(events_df)

# [Optionally] review sample observations
head(train_df, 5)
head(test_df, 5)
head(stores_df, 5)
head(oil_df, 5)
head(events_df, 5)
```


### Univariate Analysis and Data Preparation


###### Data Profiling

```{r data_profiling}
# e.g., statistical characteristics (including distribution, skewness, outliers)
```


###### Preliminary Feature Reduction (clearly n/a to Objective and Hypothesis)


###### Attribute Names (confirm/update)


###### Target and Feature Identification (preliminary)


###### Data Types (confirm/update)


###### Data Enrichment


###### Categorical Data Profiling


### Data Partitioning


### Additional Uni/Multivariate EDA and Feature Engineering/Selection


###### Missing/null values (find/impute/drop)


###### Categorical features (encoding)


###### Outliers (convert/drop)


###### Zero/Near-Zero Variances (find/address)


###### Centering/scaling (standardizing/normalizing)


###### Collinearity and Dependencies (find/address)


###### "Noisy"/duplicate data (find/convert/drop)


### Data Mining (Unsupervised)


### Modeling


###### Final Feature Selection


###### Model Setup (Selection)


###### Model Run and Evaluation (Iteration n)


###### Optimization, Tuning, Selection

