---
title: "Grocery Sales Forecast"
author:
  Christine Vu^[University of San Diego, cvu@sandiego.edu], Dave Friesen^[University of San Diego, dfriesen@sandiego.edu]
date: "12/12/2022"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output:
  html_document:
    css: "style.css"
  pdf_document: default
---

<style>
.main-container {
  max-width: 1024px;
}
</style>


### Objective and Hypothesis

#### [. . .]


```{r setup, echo = FALSE, message = FALSE}
# Load R libraries
library(e1071)
library(readr)
library(RTextTools)
library(readxl)
library(ggplot2)
library(plotly)
library(forecast)
library(dplyr)
library(tidyr)
library(zoo)

# Expand output width and minimize exp notation
options(width = 150)
options(scipen = 100)
options(digits = 1)

# Set style defaults
knitr::opts_chunk$set(class.source = "source")
knitr::opts_chunk$set(class.output = "output")
knitr::opts_chunk$set(fig.width = 10, fig.height = (10 * .45), fig.align = "center")
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NA)
```


### Data Load and Validation

```{r data_load_validation}
# Load dataset(s); assumes folder structure with data parallel to src
train_df <- read.csv("/Users/christinevu/Downloads/train.csv", header = TRUE)
test_df <- read.csv("/Users/christinevu/Downloads/test.csv", header = TRUE)
stores_df <- read.csv("/Users/christinevu/Downloads/stores.csv", header = TRUE)
oil_df <- read.csv("/Users/christinevu/Downloads/oil.csv", header = TRUE)
events_df <- read.csv("/Users/christinevu/Downloads/holidays_events.csv", header = TRUE)

# Data validation and understanding, including structure, content, and statistical characteristics covered below
```


###### Data Structure Review

```{r data_structure_review}
# Summarize base datasets
str(train_df)
str(test_df)
str(stores_df)
str(oil_df)
str(events_df)

# [Optionally] review sample observations
head(train_df, 5)
head(test_df, 5)
head(stores_df, 5)
head(oil_df, 5)
head(events_df, 5)
```


### Univariate Analysis and Data Preparation


###### Data Profiling

```{r data_univariate, echo = FALSE}
# Note this function is generic and doesn't look for more intelligent "blank" values like "no record",
#   "not available", etc.
is_blank <- function(x) {
  classof_x <- class(x)
  result <-
    !is.na(x) &
    (((classof_x == "character") & (x == "")) |
     ((classof_x %in% c("integer", "numeric")) & (x == 0)))
  return(result)
}

# Function to format percentages (only when value exists)
format_percent <- function(x) {
  result <- formatC(x * 100, digits = 0, width = 5, format = "d", zero.print = FALSE)
  if (round(x, 0.5) != 0) result <- paste(result, "%", sep = "")
  return(result)  
}

# Function to not output NaNs from third-party functions in lapply() below
nan_replace_0 <- function(x) {
  if (is.nan(x)) result <- 0 else result = x
  return(result)
}

# Function to Generate a summary of base dataset
univariate <- function(df) {
  rowcount <- nrow(df)
  ua <- do.call(rbind, lapply(df, function(x) c(
    colnames(x),
    class(x),
    format_percent(sum(is.na(x)) / rowcount),
    format_percent(sum(is_blank(x)) / rowcount),
    formatC(length(unique(na.omit(x))),
            digits = 0, width = 9, format = "d", big.mark = ",", zero.print = FALSE),
    formatC(ifelse(is.numeric(x), min(na.omit(x)), 0),
            digits = 0, width = 9, format = "f", big.mark = ",", zero.print = FALSE),
#            digits = ifelse(is.double(x), 3, 0), width = 9, format = "f", big.mark = ",", zero.print = FALSE),
    formatC(ifelse(is.numeric(x), max(na.omit(x)), 0),
            digits = 0, width = 9, format = "f", big.mark = ",", zero.print = FALSE),
#            digits = ifelse(is.double(x), 3, 0), width = 9, format = "f", big.mark = ",", zero.print = FALSE),
    formatC(ifelse(is.double(x), mean(na.omit(x)), 0),
            digits = 0, width = 9, format = "f", big.mark = ",", zero.print = FALSE),
#            digits = 3, width = 9, format = "f", big.mark = ",", zero.print = FALSE),
    formatC(ifelse(is.numeric(x), median(na.omit(x)), 0),
            digits = 0, width = 9, format = "f", big.mark = ",", zero.print = FALSE),
#            digits = ifelse(is.double(x), 3, 0), width = 9, format = "f", big.mark = ",", zero.print = FALSE),
    format(ifelse(is.numeric(x),
           ifelse(na.omit(x) < (quantile(na.omit(x), 0.25) - (1.5 * IQR(na.omit(x)))), "Yes", "No"), ""),
           justify = "centre", width = 8, format = "s"),
    format(ifelse(is.numeric(x),
           ifelse(na.omit(x) > (quantile(na.omit(x), 0.75) - (1.5 * IQR(na.omit(x)))), "Yes", "No"), ""),
           justify = "centre", width = 8, format = "s"),
    formatC(ifelse(is.numeric(x), nan_replace_0(skewness(na.omit(x))), 0),
            digits = 2, width = 8, format = "f", zero.print = FALSE),
    formatC(ifelse(is.numeric(x), nan_replace_0(kurtosis(na.omit(x))), 0),
            digits = 2, width = 8, format = "f", zero.print = FALSE))))
  colnames(ua) <- c(
    "Type",
    format("NA", justify = "right", width = 6),
    format("BlankZ", justify = "right", width = 6),
    format("Unique", justify = "right", width = 9),
    format("Min", justify = "right", width = 9),
    format("Max", justify = "right", width = 9),
    format("Mean", justify = "right", width = 9),
    format("Median", justify = "right", width = 9),
    format("Outlier<", justify = "centre", width = 8),
    format(">Outlier", justify = "centre", width = 8),
    format("Kurtosis", justify = "right", width = 8),
    format("Skewness", justify = "right", width = 8))
  row.names(ua) <- lapply(row.names(ua),
                          function(x) if (nchar(x) > 20) return(paste(substr(x, 1, 17), "...", sep = ""))
                          else return(x))
  { cat(
    "Summary Univariate Analysis (",
    formatC(rowcount, big.mark = ","), " observations)\n",
    sep = "")
    print(noquote(ua))
  }
}
```

```{r data_profiling}
# e.g., statistical characteristics (including distribution, skewness, outliers)
univariate(train_df)
univariate(test_df)
univariate(stores_df)
univariate(oil_df)
univariate(events_df)
```

```{r}
# Data summary statistics
summary(train_df)
summary(test_df)
summary(stores_df)
summary(oil_df)
summary(events_df)
```

```{r}
# Proportion by family
fam_prop <- train_df %>%
              group_by(family) %>%
              summarise(Mean_Sales = round(mean(sales, na.rm = TRUE), 2),
                        Median_Sales = round(median(sales, na.rm = TRUE), 2),
                        Maximum_Sales = max(sales),
                        IQR_Sales = IQR(sales),
                        Total_Sales = sum(sales))

fam_prop
```

```{r}
# Box plot: Mean of store sales by family
fam_plot <- ggplot(fam_prop) +
            geom_col(aes(x = family, y = Mean_Sales, fill = family)) +
            theme(axis.ticks.x = element_blank(),
                  axis.text.x = element_blank()) +
            labs(title = "Average Store Sales by Family",
                 x = "Family of Product",
                 y = "Average of Daily Sales",
                 legend = "Family of Product Sold",
                 bty = "l")

ggplotly(fam_plot)
```

```{r}
# Box plot: Total store sales by family
fam_plot2 <- ggplot(fam_prop) +
             geom_col(aes(x = family, y = Total_Sales, fill = family)) +
             theme(axis.ticks.x = element_blank(),
                   axis.text.x = element_blank()) +
             labs(title = "Total Store Sales by Family",
                  x = "Family of Product",
                  y = "Average of Daily Sales",
                  legend = "Family of Product Sold",
                  bty = "l")

ggplotly(fam_plot2)
```


###### Preliminary Feature Reduction (clearly n/a to Objective and Hypothesis)


###### Attribute Names (confirm/update)


###### Target and Feature Identification (preliminary)


###### Data Types (confirm/update)
```{r}
lapply(train_df, class)
lapply(test_df, class)
lapply(stores_df, class)
lapply(oil_df, class)
lapply(events_df, class)
```


###### Time Series Objects
```{r}
# Convert data to time series objects
train.ts <- ts(train_df$sales, start = c(1, 1), freq = 52)
test.ts <- ts(test_df$family, start = c(1, 1), freq = 52)
stores.ts <- ts(stores_df$cluster, start = c(1, 1), freq = 52)
oil.ts <- ts(oil_df$dcoilwtico, start = c(1, 1), freq = 52)
#events.ts <- ts(holiday_events$sales, start = c(1, 1), freq = 52)
```


###### Data Enrichment


###### Categorical Data Profiling


### Data Partitioning
```{r}
# Count rows
nrow(train_df)
nrow(test_df)

# Count columns
ncol(train_df)
ncol(test_df)
```


### Additional Uni/Multivariate EDA and Feature Engineering/Selection


###### Missing/null values (find/impute/drop)
```{r}
# Count missing values
sum(is.na(train_df))
sum(is.na(test_df))
sum(is.na(stores_df))
sum(is.na(oil_df))
sum(is.na(events_df))
```

```{r}
# Mean of missing oil_df data
mean(is.na(oil_df))

# Option: Drop missing records since they only account for 2% of data
# na.omit(oil_df)
```

```{r}
# oil_df2: New 'oil_df' data frame without missing values
# Foward and backward fill for missing values
oil_df2 <- na.locf(oil_df, fromLast = TRUE)

# Confirm oil_df2 has no more missing values
sum(is.na(oil_df2))
```


###### Categorical features (encoding)


###### Outliers (convert/drop)
```{r}
# Box plot: oil_df2
ggplot(oil_df2) +
  aes(x = "", y = dcoilwtico) + geom_boxplot(fill = "#0c4c8a") +
  theme_minimal() + labs(title = "Oil Box Plot", x = "", y = "Oil Price")
```

###### Zero/Near-Zero Variances (find/address)


###### Centering/scaling (standardizing/normalizing)


###### Collinearity and Dependencies (find/address)


###### "Noisy"/duplicate data (find/convert/drop)


### Data Mining (Unsupervised)


### Modeling


###### Final Feature Selection


###### Model Setup (Selection)


###### Model Run and Evaluation (Iteration n)


###### Optimization, Tuning, Selection

